A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-10 survey.tex
Revised Section IV

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration), there hasn't been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have not been explored so far, and would be a very interesting venue of future work.

@@

Playing an RTS game involves dealing with all the problems described above. A few approaches, like CAT \cite{LTW}, Darmok \cite{OntanonMSR10} or ALisp \cite{Marthi05} try to deal with the problem in a monolithic manner, by using a single AI technique. This resembles approaches to solve other games, such as Chess or Go, where a single game-tree search approach is enough to play the game at human level. However, none of those systems aims at achieving near human performance. In order to achieve human-level performance, RTS AI designers use a lot of domain knowledge in order to divide the task of playing the game into a collection of sub-problems, which can be dealt-with using individual AI techniques (as discussed in the previous section).

@@

Figure \ref{fig:bot-architecture} shows some representative examples of the integration architectures used by different bots in the AIIDE and CIG StarCraft AI competitions (see Section \ref{sec:competition}): BroodwarQ \cite{SynnaeveMicroCig11}, Nova \cite{uriarte2012kiting}, UAlbertaBot \cite{churchill2011build}, Skynet, SPAR, AIUR, and BTHAI \cite{Hagelback12}. Each box represents an individual module with a clearly defined task (only modules with a black background can send actions directly to StarCraft). Dashed arrows represent data flow, and solid arrows represent control (when a module can command another module to perform some task). For example, we can see how SPAR is divided in two sets of modules: {\em situation analysis} and {\em decision making}, the first with three modules dedicated to analyze the current situation of the game, and the later with 4 modules dedicated to exploit that information to decide what to do. We can see how the decision making aspect of SPAR is organized hierarchically, with the higher-level module ({\em strategic decision}) issuing commands to the next module ({\em tactical decision}), which sends commands to the next module ({\em action implementation}), and so on. Only the lower-level modules can send actions directly to StarCraft.

@@

By analyzing the structure of these bots, we can see that there are two main tools that can be used when designing an integration architecture:

@@

Different bots, use different combinations of these two tools. Looking back at Figure \ref{fig:bot-architecture}, we can see the following use of abstraction and divide-in-conquer in the bots:

@@

Some bots  using divide-and-conquer, assume  that each of  the modules
can act independently  and that their actions can  be executed without
interference.  BBQ,  UAlbertaBot and  AIUR, however use  an arbitrator
({\em Game Commander'} in UAlbertaBot) that makes sure that modules do
not send contradictory orders to  the same unit.  However, very little
bots handle  the problem of  how to coordinate resource  usage amongst
modules, for  instance BTHAI uses a  first-come-first-serve policy for
spending resources,  the first module  that requests resources  is the
one that gets them. Nova and Skynet are exceptions, and implement some
rudimentary prioritization based on the high level strategy. Following
available  resources and  timing,  AIUR's {\em  Spend Manager}  orders
Base,  Production   and  Construction  Managers  what   they  have  to
build/produce. It also orders to start tech research and upgrades. The
idea here is  not to allocation resources among  different managers to
satisfy a need (informally: ``I want  to do that, so I need M minerals
and G gas''), but to do the  opposite, that is, to find how the AI can
spend the available money: ``I have right now M minerals and G gas, so
how can I spend it?''.

One  interesting aspect  of the  seven bots  described above  is that,
while all  of them  - except AIUR  - are  reactive at the  lower level
(reactive  control), most  if not  all of  them, are  scripted  at the
highest level  of abstraction. BTHAI reads build  and squad formations
from  a  predefined  script,   Nova's  {\em  Strategy  Manager}  is  a
predefined finite-state machine,  BBQ's construction manager reads the
build  order from a  predefined script,  and Skynet's  {\em BuildOrder
  Manager} is basically a predefined script. Such scripts describe the
strategy  that the  bots will  use, however,  such strategy  is always
fixed.   One could see  this pre-scripting  as if  each bot  defined a
``high-level programming language''  to describe StarCraft strategies,
and   the   bots   themselves    are   just   interpreters   of   such
strategy.  Compared  to  current  approaches  for Chess  or  Go,  this
scripting  seems a  rigid and  inflexible,  but responds  to the  much
higher complexity of the  StarCraft game.  An interesting exception to
that  is  UAlbertaBot, which  uses  a  search  algorithm in  the  {\em
  Production  Manager}  to find  near-optimal  build orders.   Another
interesting case is  AIUR, that uses a {\em  Mood Manager} to randomly
pick  a mood among  six (cheese,  rush, aggressive,  defensive, macro,
fast  expand), which  will  influence the  build  order, strategy  and
tactics.  All of  them  scripted so  far,  but build  orders can  been
modified but  the Spend Manager, following  available resources.  This
mood can change during the game, regarding the opponent behavior.

In conclusion, we can see that there are two basic tools that can be used in an integration architecture: abstraction and divide-and-conquer, which are widely used by the existing StarCraft bots. For space reasons, we do not include an exhaustive comparison of the architectures of all the participating bots. Some other bots have been documented by their authors, such as SCAIL \cite{YoungSCAIL} or QUORUM \cite{young2012evolutionary}.

This section has focused on discussing the architecture of existing StarCraft bots. Let us now focus on what is the state of the art on the techniques used for each of the individual modules used by them.

>>>>

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration) or ALisp \cite{Marthi05}, there hasn't been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have not been explored so far, and would be a very interesting venue of future work.

@@

Playing an RTS game involves dealing with all the problems described above. A few approaches, like CAT \cite{LTW}, Darmok \cite{OntanonMSR10} or ALisp \cite{Marthi05} try to deal with the problem in a monolithic manner, by using a single AI technique. However, none of those systems aims at achieving near human performance. In order to achieve human-level performance, RTS AI designers use a lot of domain knowledge in order to divide the task of playing the game into a collection of sub-problems, which can be dealt-with using individual AI techniques.

@@

Figure \ref{fig:bot-architecture} shows some representative examples of the architectures used by different bots in the AIIDE and CIG StarCraft AI competitions (see Section \ref{sec:competition}): BroodwarBotQ \cite{SynnaeveMicroCig11}, Nova \cite{uriarte2012kiting}, UAlbertaBot \cite{churchill2011build}, Skynet, SPAR, AIUR, and BTHAI \cite{Hagelback12}. Each box represents an individual module with a clearly defined task (only modules with a black background can send actions directly to StarCraft). Dashed arrows represent data flow, and solid arrows represent control (when a module can command another module to perform some task). For example, we can see how SPAR is divided in two sets of modules: {\em situation analysis} and {\em decision making}. Situation analysis in SPAR has three modules dedicated to analyze the current situation of the game. Decision making in SPAR is done through four hierarchically organized modules, with the higher-level module ({\em strategic decision}) issuing commands to the next module ({\em tactical decision}), which sends commands to the next module ({\em action implementation}), and so on. Only the two lower-level modules can send actions directly to StarCraft.

@@

By analyzing the structure of these bots, we can see that there are two main tools being used in these integration architectures:

@@

Different bots use different combinations of these two tools. Looking back at Figure \ref{fig:bot-architecture}, we can see the following use of abstraction and divide-in-conquer in the bots:

@@

Some bots  using divide-and-conquer, assume  that each of  the modules
can act independently  and that their actions can  be executed without
interference.  BBQ,  UAlbertaBot and  AIUR, however use  an arbitrator
({\em Game Commander'} in UAlbertaBot) that makes sure that modules do
not send contradictory orders to  the same unit.  However, very little
bots handle  the problem of  how to coordinate resource  usage amongst
modules, for  instance BTHAI uses a  first-come-first-serve policy for
spending resources,  the first module  that requests resources  is the
one that gets them. Nova and Skynet are exceptions, and implement some
rudimentary prioritization based on the high level strategy. Following
available  resources and  timing,  AIUR's {\em  Spend Manager}  orders
Base,  Production   and  Construction  Managers  what   they  have  to
build/produce. It also orders to start tech research and upgrades. The
idea here is not to let the different managers allocate the resources they want, but to do the opposite, that is: finding how the AI can
spend the available money.

One  interesting aspect  of the  seven bots  described above  is that,
while all  of them (except AIUR) are  reactive at the  lower level
(reactive  control), most  if not  all of  them, are  scripted  at the
highest level  of abstraction. BTHAI reads build  and squad formations
from  a  predefined  script,   Nova's  {\em  Strategy  Manager}  is  a
predefined finite-state machine,  BBQ's construction manager reads the
build  order from a  predefined script,  and Skynet's  {\em BuildOrder
  Manager} is basically a predefined script. Such scripts describe the
strategy  that the  bots will  use, however,  such strategy  is always
fixed.   One could see  this pre-scripting  as if  each bot  defined a
``high-level programming language''  to describe StarCraft strategies,
and   the   bots   themselves    are   just   interpreters   of   such
strategy.  Compared  to  current  approaches  for Chess  or  Go,  this
scripting  seems a  rigid and  inflexible,  but responds  to the  much
higher complexity of the  StarCraft game.  An interesting exception to
that  is  UAlbertaBot, which  uses  a  search  algorithm in  the  {\em
  Production  Manager}  to find  near-optimal  build orders.   Another
interesting case is  AIUR, that uses a {\em  Mood Manager} to randomly
pick  a mood among  six (cheese,  rush, aggressive,  defensive, macro,
fast  expand), which  will  influence the  build  order, strategy  and
tactics. % All of  them  scripted so  far,  but build  orders can  been modified but  the Spend Manager, following  available resources.  This mood can change during the game, regarding the opponent behavior.

In conclusion, we can see that there are two basic tools that can be used in an integration architecture: abstraction and divide-and-conquer, which are widely used by the existing StarCraft bots. For space reasons, we do not include an exhaustive comparison of the architectures of all the participating bots. Some other bots have been documented by their authors, such as SCAIL \cite{YoungSCAIL} or QUORUM \cite{young2012evolutionary}.
%This section has focused on discussing the architecture of existing StarCraft bots. 
Let us now focus on their performance.
